\documentclass[letter]{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{ifthen}
\usepackage{fancyhdr}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}

%%%
% Set up the margins to use a fairly large area of the page
%%%
\oddsidemargin=.2in
\evensidemargin=.2in
\textwidth=6in
\topmargin=-.4in
\textheight=9.0in
\parskip=.07in
\parindent=0in
\pagestyle{fancy}

%%%
% Set up the header
%%%
\newcommand{\setheader}[6]{
	\lhead{{\sc #1}\\{\sc #2} ({\small \it \today})}
	\rhead{
		{\bf #3} 
		\ifthenelse{\equal{#4}{}}{}{(#4)}\\
		{\bf #5} 
		\ifthenelse{\equal{#6}{}}{}{(#6)}%
	}
}

%%%
% Set up some shortcut commands
%%%
\newcommand{\R}{\mathbb{R}}
\renewcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Proj}{\mathrm{proj}}
\newcommand{\Perp}{\mathrm{perp}}
\newcommand{\proj}{\mathrm{proj}}
\newcommand{\Span}{\mathrm{span}}
\newcommand{\Null}{\mathrm{null}}
\newcommand{\Rank}{\mathrm{rank}}
\newcommand{\mat}[1]{\begin{bmatrix}#1\end{bmatrix}}
\renewcommand{\d}{\mathrm{d}}

%%%
% This is where the body of the document goes
%%%
\begin{document}
	\setheader{Math 281-3}{Homework 4}{Due Thursday, April 20}{}{}{}


	For this homework, you will be using \emph{Matlab} (or the free and open-source version of
	\emph{Matlab} called \emph{Octave}).  Include a clearly-labeled printout
	of the work you did in Matlab for each problem.
	
	You can read through \\
	\centerline{\url{http://web.gps.caltech.edu/classes/ge11d/doc/matlab_Resource_Seminar.pdf}}\\
	for a short Matlab introduction.
	For additional documentation, see the Matlab website \\\centerline{\url{http://www.mathworks.com/help/matlab/index.html}}

	\textbf{Quick Tips:}

	\vspace{-1em}
	\begin{itemize}
		\item You can supress Matlab output by adding a `\texttt{;}'
		to the end of your Matlab command (this is useful for
		Matlab commands executed in loops).

		\item You can save plots/figures from the figure's dialog
		by selecting `Save as' and then choosing \texttt{png}
		or \texttt{jpg} in the file-types dropdown.

		\item You can include Matlab code in \LaTeX\ by copying
		and pasting it between \texttt{\textbackslash
		begin\{verbatim\}} \ldots\texttt{\textbackslash
		end\{verbatim\}}.
	\end{itemize}

	\begin{enumerate}
		\item Find the general solution to the system of equations corresponding to each augmented matrix.
			\begin{enumerate}
				\item \[\left[\begin{array}{ccccc|c}
				       1&    11&   -22&     0&     3&1 \\
				    5244&  1542& -3084&   576& 16308&2\\
				     108&  1188& -2376&     0&   324&3\\
				     264&   465&  -930&    25&   817&4\\
				    2166&  3729& -7458&   206&  6704&5
					\end{array}\right]\]
				\item \[\left[\begin{array}{ccccc|c}
						    7&  3&  2&  3&  7&  1 \\
						    9&  3&  2&  1& 10&  2\\
						    7&  5&  2&  2&  8&  3\\
						    4& 10&  3&  1& 10&  4\\
						    4&  6&  1&  3&  9&  5
				    \end{array}\right]
					\]
			\end{enumerate}
		
		\item Although the \texttt{rref} command will row-reduce for you, Matlab makes it easy
			to row reduce ``by hand.''  If \texttt{A} is a matrix, the row operation 
			$R_2\mapsto R_2-6R_1$ can be performed with the command \texttt{A(2,:)=A(2,:)-6*A(1,:)}
			and other row operations can be performed similarly.

			\begin{enumerate}
				\item Consider the $3\times 3$ matrix $A=\mat{1&4&0\\1&5&0\\0&0&0}$ and $B=\mat{1&0&0\\0&1&0\\0&0&1}$.
					You only need to perform two row operations to fully reduce $A$.  Call these two
					operations $R^{(1)}$ and $R^{(2)}$. Use Matlab to perform those
					row operations on $A$; also, perform the same row operations on $B$ (effectively, un-row
					reducing $B$) and call the result $B'$.
				\item Let $B_1=R^{(1)}(B)$ and $B_2=R^{(2)}(B)$ be the matrices obtained from $B$ by applying
					the two row operations in isolation (in Matlab, you can name variables
					\texttt{B1} and \texttt{B2} if you like).  What happens when you perform the matrix product
					$B_1A$?  How about $B_2A$?  And $B_2B_1A$?  Finally, compute $B'A$.  Explain your results.
				\item Let $X=\mat{2&1&3\\2&1&3\\3&2&4}$.  Find a matrix $Y$ so that $YX=\mathrm{rref}(X)$.
			\end{enumerate}
		\item A loop in Matlab/Octave is written with the 
	\begin{verbatim}
for i=1:n,
    [loop content]
end
	\end{verbatim}
	syntax.

	For example, if we want the variable $x$ to be an accumulated sum of squares from $1^2$ to $15^2$ 
	(that is $x=1^2+2^2+\cdots+15^2$), we could write
	\begin{verbatim}
x=0;
for i=1:15,
    x = x + i^2;
end
	\end{verbatim}
	After executing this code $x$ will be the value $1240$.  The code works as follows: first we assign the value
	$0$ to $x$.  Then we enter the loop.  In the first iteration, when $i$ is 1, the new value of $x$ will be the current value,
	0, plus $1^2$.  The next time through the loop, $i$ is 2 and so we assign $x$ to be the current value, $1^2$, plus
	$2^2$, etc.. Once we have looped through with the value of $i$ being 15, we stop.

	If statements are written with the 
	\begin{verbatim}
if [var]==[val],
    [if content]
end
	\end{verbatim}
	syntax (note the double equals). 

	For example, if we wanted $x$ to be the sum of the squares of only the even numbers between 1 and 15, we might write:
	\begin{verbatim}
x=0;
for i=1:15,
    if round(i/2) == i/2,
        x = x + i^2;
    end
end
	\end{verbatim}

	Here, {\tt round(i/2)==i/2} is true when {\tt i/2} has no decimal places and so it is even.  Thus, when that happens
	(and only when that happens) we add the value $i^2$ to $x$.


	{\bf Question}:
	\begin{enumerate}
		\item The command \texttt{rand(3,1)} will generate a random $3\times 1$ column vector.  Find out what percentage
		of random triples of vectors $\{\vec a,\vec b,\vec c\}$ are linearly independent.  Sample at least 1000
		random triples to gather a statistic.

		Consider your result and come up with an explanation of why you got the percentage you did.

		{\bf Tips}: Build a loop to test this 1000 times, but start out
		with a much smaller loop;
		use `\texttt{;}' inside your loop so you don't get a bunch of unhelpful information printed to the screen;
		and use Matlab's built-in commands to help you like \texttt{rref} to row reduce, \texttt{rank} to find the
		rank, or \texttt{rand(3,3)} to make a $3\times 3$ matrix of random numbers.

		\item The command \texttt{round(rand(3,1))} will generate a random $3\times 1$ column vector of just zeros
			and ones.  Find out what percentage
			of random triples  $\{\vec a,\vec b,\vec c\}$ of zero-one vectors are linearly independent.  Sample at least 1000
			random triples to gather a statistic.

		Explain your result.  Why is it different or the same as the previous part?  (Hint: it may be useful
		to think of random zero-one vectors as lying on the vertices of the unit cube.)

	\end{enumerate}


	\end{enumerate}

\end{document}
